---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
```

```{r}
library(tidyverse)
```

```{r}
library(dplyr)
```

```{r}
decathlon_clean <- read_rds("decathlon_data_cleaned.rds")
decathlon_clean
```


1 MVP
1.1 Task 1 - Decathlon Data

1.1.2 Analysis questions
1. Who had the longest long jump seen in the data?

```{r}
long_jump <- select(decathlon_clean, "athlete", "long_jump")
head(long_jump,1)
long_jump
```

```{r}
long_jumpq1 <- decathlon_clean %>%
  select("athlete", "long_jump",) %>%
  arrange(desc("long_jump")) %>%
  head(1)
long_jumpq1
```


2. What was the average 100m time in each competition?


```{r}
Decastar_average_time <- filter(decathlon_clean, competition == "Decastar")
mean(Decastar_average_time$x100m)
Decastar_average_time
```

```{r}
OlympicG_average_time <- filter(decathlon_clean, competition == "OlympicG")
mean(OlympicG_average_time$x100m)
OlympicG_average_time

```


1.3 Who had the highest total points across both competitions?
```{r}
total_points_q4 <- decathlon_clean %>%
  select("athlete", "competition", "points") %>%
  filter(competition == "OlympicG") %>%
  arrange(desc(points)) %>%
  head(1)
total_points_q4
```

```{r}
total_points <- decathlon_clean %>%
  select("athlete", "competition", "points") %>%
  filter(competition == "Decastar") %>%
  arrange(desc(points)) %>%
  head(1)
total_points

```


1.4 What was the shot-put scores for the top three competitors in each competition?
```{r}
shot_put_scores_q4 <- decathlon_clean %>%
  select("athlete", "shot_put", "points", "competition") %>%
  filter(competition == "OlympicG") %>%
  arrange(desc(shot_put)) %>%
  head(3)
shot_put_scores_q4

```

```{r}
shot_put_scores <- decathlon_clean %>%
  select("athlete", "shot_put", "points", "competition") %>%
  filter(competition == "Decastar") %>%
  arrange(desc(shot_put)) %>%
  head(3)
shot_put_scores

```

1.5 What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
athlete_400_q5 <- decathlon_clean %>%
  select("competition") %>%
  filter("x400m" > 50) %>%
  group_by(competition) %>%
  summarise(average) 
athlete_400_q5

```

```{r}
athlete_400 <- decathlon_clean %>%
  select("competition") %>%
  filter("x400m" < 50) %>%
  group_by(competition) %>%
  summarise(average) 
athlete_400
```






